---
runtime: shiny
title: "BiotaMetaller"
author: "Martin Sköld"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(readxl)
library(ggmap)
library(rnaturalearth)
library(rnaturalearthdata)
library(lmtest)
library(cowplot)
library(rgeos)
theme_set(theme_minimal())
```

## Data wrangling (data downloaded from  [ivl](https://dvsb.ivl.se/SokResMETALS?m=2&from=1849&tom=2016&laen=AA) 2019-05-21)

```{r}
BiotaMetaller <- read_excel("BiotaMetaller.xlsx", 
                            skip = 1, guess_max = 5000) %>% 
  select(Station, Longitude, Latitude, Provdatum, ends_with("DW")) %>% 
  mutate_if(is.logical, as.double) %>% 
  mutate(Year = format(Provdatum, "%Y") %>% as.numeric()) %>% 
  gather(key = Metal, value = Concentration, ends_with("DW")) %>% 
  mutate(Metal = str_sub(Metal, 1, 2)) %>% 
  filter(!is.na(Concentration))
glimpse(BiotaMetaller)
```


## Webb-app (time series)

```{r, warning=FALSE, message=FALSE}
shinyApp(options = list(height = 500),
  ui <- fluidPage(
    fluidRow(
      column(4,
             wellPanel(
               selectInput("Metal", "Metal",
                           choices = BiotaMetaller %>% pull(Metal) %>% unique()
               ),
               selectInput("Stations", "Station", 
                           choices = BiotaMetaller %>% pull(Station) %>% unique(), 
                           selected = "Abiskojaure",
                           multiple = TRUE),
               sliderInput("Years", "Years", 1965, 2015, c(2006, 2015), sep = ""),
               selectInput("Trend", "Trend", choices = c("Log-linear", "Smooth")),
               checkboxInput("Interval", "Confidence intervall (95%)")
             )
      ),
      column(8,
             plotOutput("plot"))
    )
  ),
  server <- function(input, output) {
    plot_data <- reactive({
      BiotaMetaller %>% 
        filter(Metal == input$Metal, 
               Station %in% input$Stations,
               Year %in% input$Years[1]:input$Years[2])
    })
    output$plot <- renderPlot(
      plot_data() %>% 
        ggplot(aes(x = Provdatum, y = log(Concentration), color = Station)) + 
        geom_point(alpha = .5) + 
        geom_smooth(se = input$Interval, 
                    method = ifelse(input$Trend == "Log-linear", "lm", "loess"))
    )
  }
  
)
```


## Trends summary

```{r}
BiotaMetaller %>% 
  filter(Year > 2005) %>% 
  filter(Concentration > 0) %>% 
  group_by(Station, Year, Metal) %>% 
  summarise(mean_Conc = mean(log(Concentration), na.rm = TRUE)) %>% 
  group_by(Station, Metal) %>% 
  nest() %>% 
  filter(map(data, nrow) > 5) %>%
  mutate(model = map(data, ~lm(mean_Conc ~ Year, .x)), slope = map_dbl(model, ~coef(.x) %>% .[2])) %>% 
  ggplot() + 
  geom_histogram(aes(x = (exp(slope)-1)*100)) + 
  facet_wrap(~ Metal) +  
  geom_vline(xintercept = 0, color = "red") + 
  xlab("Average yearly change (%) 2006-2015 (stations with more than 5 samplings)") + 
  xlim(c(-30, 30))
```

## Serial dependence test, P-values

```{r}
BiotaMetaller %>% 
  filter(Year > 1965) %>% 
  filter(Concentration > 0) %>% 
  filter(!is.na(log(Concentration))) %>% 
  group_by(Station, Year, Metal) %>% 
  summarise(m = mean(log(Concentration), na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(Station, Metal) %>% nest() %>% 
  filter(map(data, nrow) > 5) %>% 
  mutate(model = map(data, ~lm(m ~ Year,.x)), slope = map_dbl(model, ~coef(.x) %>% .[2]),
         dw_p = map_dbl(data, ~dwtest(m ~ Year, data = .x) %>% .["p.value"] %>% unlist())) -> model_df2
model_df2 %>% ggplot() + geom_histogram(aes(x = dw_p), binwidth = 0.05) + facet_wrap(~ Metal) + theme_minimal() + xlab("P-värde (Durbin-Watson test)")


## Webb-app (map)

```{r}
world_map <- ne_countries(scale = "medium", returnclass = "sf")

shinyApp(options = list(height = 800),
  ui <- fluidPage(
    fluidRow(
      column(2,
             wellPanel(
               selectInput("Metal", "Metal",
                           choices = BiotaMetaller %>% pull(Metal) %>% unique()
               ),
               sliderInput("Year", "Year", 1965, 2015, 2015, sep = "")
             )), column(10,
                        plotOutput(outputId = "mymap", height = "510px")
             )
    )),
  
  server <- function(input, output){
    mapdata <- reactive({
      BiotaMetaller %>% 
        filter(Year == input$Year,
               Metal == input$Metal,
               Concentration > 0) %>% 
        group_by(Station, Longitude, Latitude) %>% 
        summarise(mean_Conc = mean(log(Concentration), na.rm = TRUE)) %>% 
        ungroup()
    })
    
    output$mymap <- renderPlot({
      plot_grid(ggplot(data = world_map) + 
        geom_sf() + 
        coord_sf(xlim = c(8, 25), ylim = c(54, 70), expand = FALSE) +
        geom_point(data = mapdata(), aes(x = Longitude, y = Latitude, color = mean_Conc, size = mean_Conc), alpha = .8) + 
          theme(legend.position = "none"),
        ggplot(mapdata(), aes(x = fct_reorder(Station, Latitude), y = exp(mean_Conc))) + geom_col()+  coord_flip() +xlab("") +ylab("Concentration"))})
  }
)
```

